---
title: "Mini-Project"
author: "krishnakanth-20233"
date: "3/5/2021"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries

```{r}
library(plyr)
library(tidyverse)
library(gridExtra)
library(gmodels)
library(corrplot)
library(MASS)
library(ISLR)
library(caret)
library(leaps)
library(car)
```

# Data setup

```{r}
data <- read_csv("heart.csv") # reading dataset
data <- na.omit(data)
df<- data # copy of data
head(data) 
dim(data)
```


# Data types

```{r}
glimpse(data) # Datatype and values of the variable
str(data) # Datatype with explanation of that variable and their values
```

# conversion to factor

```{r}
data$sex <- as.factor(data$sex)
data$cp <- as.factor(data$cp)
data$fbs <- as.factor(data$fbs)
data$restecg <- as.factor(data$restecg)
data$exang <- as.factor(data$exang)
data$slope <- as.factor(data$slope)
data$ca <- as.factor(data$ca)
data$thal <- as.factor(data$thal)
data$target <- as.factor(data$target)
str(data)
```

#  Basic Statistics and Exploratory Data Analysis
 
## Summary

```{r}
summary(data)
```

## Variance

```{r}
var(data[,-c(2,3,6,7,9,11,12,13,14)])
```

## correlation matrix

```{r}
cor(data[,-c(2,3,6,7,9,11,12,13,14)])
```

## Cross tables

```{r}
CrossTable(data$target,data$sex)
```

## QQ plots

```{r}
grid.arrange(
  ggplot(data,aes(sample=age))+geom_qq(col="darkslategray3")+ggtitle("Age")+geom_qq_line(),
  ggplot(data,aes(sample=trestbps))+geom_qq(col="deeppink2")+ggtitle("Blood pressure")+geom_qq_line() ,
  ggplot(data,aes(sample=chol))+geom_qq(col="indianred2")+ggtitle("cholesterol")+geom_qq_line() ,
  ggplot(data,aes(sample=thalach))+geom_qq(col="mediumpurple1")+ggtitle("Max Heart rate")+geom_qq_line(),
  ggplot(data,aes(sample=oldpeak))+geom_qq(col="olivedrab")+ggtitle("Depression")+geom_qq_line() ,
nrow = 3,top = "QQ plots")
```

## Density curves

```{r}
grid.arrange(
  ggplot(data=data, aes(age)) + geom_density(color='chocolate3') ,
  ggplot(data=data, aes(trestbps)) + geom_density(color='chocolate3') ,
  ggplot(data=data, aes(chol)) + geom_density(color='chocolate3') , 
  ggplot(data=data, aes(thalach)) + geom_density(color='chocolate3') ,
  ggplot(data=data, aes(oldpeak)) + geom_density(color='chocolate3')+xlim(c(-2,6)) ,
  nrow=3, top = 'Density curves')
```

## shapiro test

```{r}
sp<-Map(function(x)cbind(shapiro.test(x)$statistic,shapiro.test(x)$p.value),data[,-c(2,3,6,7,9,11,12,13,14)])
output<-ldply(sp)
names(output)<-c("var","W","p.value")
flag<-0
for(i in 1:5)
{
  if(output[i,]$p.value > 0.05)
  {
    if(flag == 0)
      print("with 95 % confidence interval Normal varibles:")
    print(output[i,]$var)
    flag <- flag+1
  }
}
if(flag == 0)
  print("No variable is Normal in the data with 5% signifinace level")
```


## Histograms 

```{r}
grid.arrange(
  ggplot(data=data, aes(age)) + geom_histogram(binwidth = 1.5,color='chocolate3') ,
  ggplot(data=data, aes(trestbps)) + geom_histogram(binwidth = 3,color='chocolate3') ,
  ggplot(data=data, aes(chol)) + geom_histogram(binwidth = 6,color='chocolate3') , 
  ggplot(data=data, aes(thalach)) + geom_histogram(binwidth = 3,color='chocolate3') ,
  ggplot(data=data, aes(oldpeak)) + geom_histogram(binwidth = 0.3,color='chocolate3') ,
  nrow=3, top = 'Histograms')
```

## Bar plots

```{r}
grid.arrange(
  ggplot(data=data, aes(sex)) + geom_bar(aes(fill=sex))+scale_fill_manual(values = c("brown1","mediumslateblue")),
  ggplot(data=data, aes(target)) + geom_bar(aes(fill=target))+scale_fill_manual(values = c("lawngreen","mediumpurple")),
  ggplot(data=data, aes(exang)) + geom_bar(aes(fill=exang))+scale_fill_manual(values = c("brown1","lightgoldenrod1")),
  ggplot(data=data, aes(fbs)) + geom_bar(aes(fill=fbs))+scale_fill_manual(values = c("mediumslateblue","lightgoldenrod1")),
  nrow=2, top = 'Barplots')
```

## Pie charts

```{r}
grid.arrange(
  ggplot(data) + geom_col(aes(x = 1, y = length(data) , fill = cp),position="fill") + coord_polar(theta = "y"),
  ggplot(data) + geom_col(aes(x = 1, y = length(data) , fill = slope),position="fill") + coord_polar(theta = "y"),
  ggplot(data) + geom_col(aes(x = 1, y = length(data) , fill = restecg),position="fill") + coord_polar(theta = "y"),
  ggplot(data) + geom_col(aes(x = 1, y = length(data) , fill = ca),position="fill") + coord_polar(theta = "y"),
  ggplot(data) + geom_col(aes(x = 1, y = length(data) , fill = thal),position="fill") + coord_polar(theta = "y"),
nrow=3,top="Pie chart")
```

## Boxplot

```{r}
grid.arrange(
  ggplot(data=data, aes(age)) + geom_boxplot(fill='olivedrab3'),
  ggplot(data=data, aes(trestbps)) + geom_boxplot(fill='dodgerblue3'),
  ggplot(data=data, aes(chol)) + geom_boxplot(fill='darksalmon'),
  ggplot(data=data, aes(thalach)) + geom_boxplot(fill='lightgoldenrod1'),
  ggplot(data=data, aes(oldpeak)) + geom_boxplot(fill='brown1'),
  nrow=3, top = 'Boxplots')
```

## scatterplot

```{r}
grid.arrange(
  ggplot(data=data, aes(age,trestbps)) + geom_point(col="deepskyblue1"),
  ggplot(data=data, aes(chol,age)) + geom_point(col="darkorchid3"),
  ggplot(data=data, aes(thalach,chol)) + geom_point(col="deeppink3"),
  ggplot(data=data, aes(oldpeak,thalach)) + geom_point(col="green4"),
  ggplot(data=data, aes(thalach,trestbps)) + geom_point(col="gray42"),
  ggplot(data=data, aes(chol,oldpeak)) + geom_point(col="firebrick3")+ylim(c(-1,6)),
  nrow=3, top = 'scatter plots')
```

## Corelation plot

```{r}
corrplot(cor(data[,-c(2,3,6,7,9,11,12,13,14)]),method = "number")
```

# Splitting data

```{r}
set.seed(8)

index = sample(nrow(df), 250)
train = df[index, ]
test = df[-index, ]

nrow(train)
nrow(test)
glimpse(train)
```


----------------------------------------------------------------------------------------------------------------------

# Linear models
## Linear model1: thalach ~.
### summary

```{r}
model1 <- lm(thalach ~.,train)
summary(model1)
```

### plot

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(model1)
```

## Linear model2: thalach ~age+exang+slope+cp+chol+target+trestbps
### summary

```{r}
model2 <- lm(thalach ~ age+exang+slope+cp+chol+target+trestbps,train)
summary(model2)
```

### plot

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(model2)
```

## Linear model3: thalach ~age+exang+slope+target+trestbps
### summary

```{r}
model3 <- lm(thalach ~ age+exang+slope+target+trestbps,train)
summary(model3)
```

### plot

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(model3)
```

### Shapiro test for model3

```{r}
shapiro.test(model3$residuals)$p.value
```

## Normalizing model3

```{r}
norm <- boxcox(thalach ~ age+exang+slope+target+trestbps,data=train)
lambda <- norm$x[which.max(norm$y)]
thalachn <- ((train$thalach)^(lambda)-1)/(lambda)
```

```{r}
shapiro.test(thalachn)
```

## Linear model after normalization
### summary

```{r}
linear_model <- lm(thalachn ~age+exang+slope+target+trestbps,train)
summary(linear_model)
```

### plot

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(linear_model)
```

### Shapiro test for model3

```{r}
shapiro.test(linear_model$residuals)$p.value
```

## Prediction of Linear model

```{r}
Pred <- predict(linear_model,test,interval = "confidence",level=0.9)
Pred <- (lambda*Pred+1)**(1/lambda)

actuals <- data.frame(cbind(actuals=test$thalach, predicteds=Pred))
head(actuals,n=10)
```



----------------------------------------------------------------------------------------------------------------------

# Logistic models

## Logistic Model1 :target ~ .
### summary

```{r}
logistic_model1 = glm(target ~ ., data = train, family = "binomial")
summary(logistic_model1)
```

### predictions on train data and its accuracy

```{r}
pred <-  predict(logistic_model1, type = "response")
model_pred <- ifelse(pred>.5, "1", "0")
mean(train$target==model_pred)
```

### confusion matrix

```{r}
mat = table(Predicted=model_pred,actual=train$target)
mat

con_mat = confusionMatrix(mat, positive="1",threshold = 0.5)
con_mat
```


## Logistic Model2 :target ~ oldpeak+age+thal+cp+ca+sex+exang
### summary

```{r}
logistic_model2 = glm(target ~ oldpeak+age+thal+cp+ca+sex+exang, data = train, family = "binomial")
summary(logistic_model2)
```

### predictions on train data and its accuracy

```{r}
pred <-  predict(logistic_model2, type = "response")
model_pred <- ifelse(pred>.5, "1", "0")
mean(train$target==model_pred)
```

### confusion matrix

```{r}
mat = table(Predicted=model_pred,actual=train$target)
mat

con_mat = confusionMatrix(mat, positive="1",threshold = 0.5)
con_mat
```

## Function for predection on test data

```{r}
logistic_pred = function(model, data, pos = 1, neg = 0, cut = 0.5) 
  {
    prob = predict(model, newdata = data, type = "response")
    ifelse(prob > cut, pos, neg)
  }
```

## Prediction on test data for logistic_model2

```{r}
test_pred = logistic_pred(logistic_model2, data = test, pos = 1, neg = 0, cut = 0.5)
test_tab = table(predicted = test_pred, actual = test$target)
test_con_mat = confusionMatrix(test_tab, positive = "1")
test_con_mat
```

----------------------------------------------------------------------------------------------------------------------

# One way Anova
## Summary

```{r}
one_way_anova <- aov(age ~ target,data)
summary(one_way_anova)
round(7.52*exp(-05),6)
```

## plots

```{r}
par(mfrow=c(2,2),mar=c(4,4,3,2))
plot(one_way_anova)
```

## shapiro test

```{r}
shapiro.test(one_way_anova$residuals)
```

## Variance test

```{r}
leveneTest(age ~ target, data, center ="mean")
```

## one way anova with assumption of equal variance

```{r}
oneway.test(age ~ target,data,var.equal = T)
```

----------------------------------------------------------------------------------------------------------------------

# Two way anova
## Two way anova test (additive model).

Additive model makes an assumption that the two factor variables are independent
```{r}
two_way_anova <- aov(age ~ cp+target,data)
summary(two_way_anova)
```

## Plot

```{r}
par(mfrow=c(2,2),mar=c(4,4,3,2))
plot(two_way_anova)
```

## checking for independence using interaction terms

```{r}
two_way_anova <- aov(age ~ cp*target,data)
summary(two_way_anova)
```

## shapiro test

```{r}
shapiro.test(two_way_anova$residuals)
```

## Variance test

```{r}
leveneTest(age ~ target*cp, data, center ="mean")
```